Plug-and-play dual-tree algorithm runtime analysis

### 2014-12-02 bounds-files/sec03-approx-kde.tex
Some changes and a better range search proof using Pari's assumptions.
Propagate changes to Lemma 1 to the document.

    \medskip
    \STATE $f_p(p_q) \gets f_p(p_q) + \mathcal{K}(p_q, p_r)$
\ref{alg:kde-bc-update}
    \RETURN $\mathcal{K}(p_q, p_r)$
  \end{algorithmic}

@@

\begin{thm}
Assume that $\mathcal{K}(\cdot, \cdot)$ is a kernel satisfying the assumptions
of the previous section.  Then, given a query set $S_q$ and a reference set
$S_r$ with expansion constant $c_r$, and using the approximate kernel density
estimation \texttt{BaseCase()} and \texttt{Score()} as given in Algorithms
\ref{alg:kde_base_case} and \ref{alg:kde_score}, respectively, with the
traversal given in Algorithm \ref{alg:cover-tree-dual}, the running time of
approximate kernel density estimation for some $\epsilon$ error parameter is
bounded by
%\begin{equation}
$O(c_r^{8 + \log_2 \zeta} (N + \theta))$
%\end{equation}
where $\zeta = -\mathcal{K}'(h) \mathcal{K}^{-1}(\epsilon) \epsilon^{-1}$ and
$\theta$ as in Theorem \ref{thm:ct-runtime}.

@@

Therefore, $R^* = \emptyset$ when
%\begin{eqnarray}
$2^{s_r^{\max} + 2} \le -\epsilon / \mathcal{K}'(h)$, or when
%\begin{equation}
$s_r^{\max} \le \log_2( -\epsilon / \mathcal{K}'(h) ) - 2$.
%\end{equation}
%
Consider, then, the case when $s_r^{\max} > \log_2( -\epsilon /
\mathcal{K}'(h) ) - 2$.  It is trivial to see that for any $\mathscr{N}_r \in
R^*$, $\mathcal{K}(d_{\min}(\mathscr{N}_q, \mathscr{N}_r)) > \epsilon$;
alternately stated, this means that
%\begin{equation}
$d(p_q, p_r) < \mathcal{K}^{-1}(\epsilon) + 2^{s_r^{\max} + 1}$.
%\end{equation}
Therefore, bounding the number of points in the set
%\begin{equation}
$B_{S_r}(p_q, \mathcal{K}^{-1}(\epsilon) + 2^{s_r^{\max} + 1}) \cap
C_{s_r^{\max}}$
%\end{equation}
is sufficient to bound $|R^*|$.  For notational convenience, define $\omega = \mathcal{K}^{-1}(\epsilon) /
2^{s_r^{\max} + 1} + 1$.  Using Lemma \ref{lem:packing} with $\delta =
2^{s_r^{\max}}$ and $\rho = 2 \omega$ gives $|R^*| = c_r^{3 + \log_2 \omega}$.

@@

The value $\omega$ is maximized when $s_r^{\max}$ is minimized.  Using the lower
bound on $s_r^{\max}$, $\omega$ is bounded as
%
%\begin{equation}
%$\omega = \frac{\mathcal{K}^{-1}(\epsilon)}{2^{s_r^{\max} + 1} + 1} \le
%\frac{\mathcal{K}^{-1}(\epsilon)}{2^{s_r^{\max} + 1}} = -2 \mathcal{K}'(h)
%\mathcal{K}^{-1}(\epsilon) \epsilon^{-1}.
%\end{equation}
$\omega = -2 \mathcal{K}'(h) \mathcal{K}^{-1}(\epsilon) \epsilon^{-1}$.
%
Finally, with $\zeta = -\mathcal{K}'(h) \mathcal{K}^{-1}(\epsilon)
\epsilon^{-1}$, we are able to conclude that $|R^*| \le c_r^{3 + \log_2 (2
\zeta)} = c_r^{4 + \log_2 \zeta}$.  Therefore, the entire dual-tree traversal
takes $O(c_r^{8 + \log_2 \zeta} (N + \theta))$ time.

@@

Other kernels do not exhibit this effect.  Consider the Laplacian kernel:

>>>>

    \medskip
    \STATE $f_p(p_q) \gets f_p(p_q) + \mathcal{K}(p_q, p_r)$
\label{alg:kde-bc-update}
    \RETURN $\mathcal{K}(p_q, p_r)$
  \end{algorithmic}

@@

\begin{thm}
Assume that $\mathcal{K}(\cdot, \cdot)$ is a kernel satisfying the assumptions
of the previous section.  Then, given a query set $S_q$ and a reference set
$S_r$ with expansion constant $c_r$, and using the approximate kernel density
estimation \texttt{BaseCase()} and \texttt{Score()} as given in Algorithms
\ref{alg:kde_base_case} and \ref{alg:kde_score}, respectively, with the
traversal given in Algorithm \ref{alg:cover-tree-dual}, the running time of
approximate kernel density estimation for some $\epsilon$ error parameter is
bounded by
%\begin{equation}
$O(c_r^{8 + \lceil \log_2 \zeta \rceil} (N + \theta))$
%\end{equation}
where $\zeta = -\mathcal{K}'(h) \mathcal{K}^{-1}(\epsilon) \epsilon^{-1}$ and
$\theta$ as in Theorem \ref{thm:ct-runtime}.

@@

Therefore, $R^* = \emptyset$ when
%\begin{eqnarray}
$2^{s_r^{\max} + 2} \le -\epsilon / \mathcal{K}'(h)$, or when
%\begin{equation}
$s_r^{\max} \le \log_2( -\epsilon / \mathcal{K}'(h) ) - 2$.
%\end{equation}
%
Consider, then, the case when $s_r^{\max} > \log_2( -\epsilon /
\mathcal{K}'(h) ) - 2$.  It is trivial to see that for any $\mathscr{N}_r \in
R^*$, $\mathcal{K}(d_{\min}(\mathscr{N}_q, \mathscr{N}_r)) > \epsilon$;
alternately stated, this means that
%\begin{equation}
$d(p_q, p_r) < \mathcal{K}^{-1}(\epsilon) + 2^{s_r^{\max} + 1}$.
%\end{equation}
Therefore, bounding the number of points in the set
%\begin{equation}
$B_{S_r}(p_q, \mathcal{K}^{-1}(\epsilon) + 2^{s_r^{\max} + 1}) \cap
C_{s_r^{\max}}$
%\end{equation}
is sufficient to bound $|R^*|$.  For notational convenience, define $\omega = \mathcal{K}^{-1}(\epsilon) /
2^{s_r^{\max} + 1} + 1$.  Using Lemma \ref{lem:packing} with $\delta =
2^{s_r^{\max}}$ and $\rho = 2 \omega$ gives $|R^*| = c_r^{3 + \lceil \log_2
\omega \rceil}$.

@@

The value $\omega$ is maximized when $s_r^{\max}$ is minimized.  Using the lower
bound on $s_r^{\max}$, $\omega$ is bounded as
%
%\begin{equation}
%$\omega = \frac{\mathcal{K}^{-1}(\epsilon)}{2^{s_r^{\max} + 1} + 1} \le
%\frac{\mathcal{K}^{-1}(\epsilon)}{2^{s_r^{\max} + 1}} = -2 \mathcal{K}'(h)
%\mathcal{K}^{-1}(\epsilon) \epsilon^{-1}.
%\end{equation}
$\omega = -2 \mathcal{K}'(h) \mathcal{K}^{-1}(\epsilon) \epsilon^{-1}$.
%
Finally, with $\zeta = -\mathcal{K}'(h) \mathcal{K}^{-1}(\epsilon)
\epsilon^{-1}$, we are able to conclude that $|R^*| \le c_r^{3 + \lceil \log_2
(2 \zeta ) \rceil} = c_r^{4 + \lceil \log_2 \zeta \rceil}$.  Therefore, the
entire dual-tree traversal takes $O(c_r^{8 + \lceil \log_2 \zeta \rceil} (N +
\theta))$ time.

@@

Next, consider the Laplacian kernel:
$\mathcal{K}_l(d) = \exp(-d / \sigma)$.  For this kernel, $h = 0$ (that is, the
kernel is always convex), so then $\mathcal{K}'_l(h) = \sigma^{-1}$.  Simple
algebraic manipulation gives $\mathcal{K}^{-1}_l(\epsilon) = -\sigma \ln
\epsilon$, resulting in $\zeta = -\mathcal{K}'_l(h) \mathcal{K}^{-1}_l(\epsilon)
\epsilon^{-1} = \epsilon^{-1} \ln \epsilon$.  So both the Laplacian and Gaussian
kernels do not exhibit dependence on the bandwidth.
